\Large Generative Dynamics of Supreme Court Citations

### 2020-04-01 Tex/paper.tex
started writing section 4.1.2 Dependence Terms

We include a dependence term to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, popularity, and activity. We expect that the coefficients associated with each of these dependence effects will be positive. The dependence effect included in the model to account for transitivity is the count of the number of triangles in $C_{\leq t}$. The log odds interpretation of the triangle statistic is that  $\theta$ gives the change in the log odds of case $i$ citing case $j$ associated  with each additional triangle that would be created by case $i$ citing case $j$ (e.g., if opinions in cases $i$ and $j$ cite one common case $k$, the log odds that case $i$ cites case $j$ change by theta relative to the configuration in which cases $i$ and $j$ do not cite any of the same cases).

The statistic included in the model for reciprocity counts the number of mutual dyads (i.e., dyads in which cases $i$ cites case $j$, and case $j$ cites case $i$) in $C_{\leq t}$. In practice, this is the number of mutual dyads in $C_{\leq t}$, since mutual dyads must emerge among opinions that were written in the same term.The log odds interpretation of the reciprocity term is that $\theta$ gives the change in the log odds of a tie

>>>>

We include a dependence term to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, popularity, and activity. We expect that the coefficients associated with each of these dependence effects will be positive.

The common dependence effects for these four dynamics are the number of triangles (case $i$ cites cases $j$ and $k$, and case $j$ cites case $k$), the numnber of mutual dyads (case $i$ cites $j$ and vice versa), the numnber of out-2-stars (case $i$ cites $j$ and $k$), and the number of in-2-stars (case $i$ is being cited by cases $j$ and $k$) in $C_{\leq t}$. Unfortunately, adding the triangle, out-2-star, or in-2-star statistic causes model degeneracy \citep{Handcock.2003}. Model degeneracy occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network. In this scenario, the simulated networks are too different from the observed network, making the underlying distribution defined by the model extremly unrealistc.

The statistic included in the model for reciprocity counts the number of mutual dyads (i.e., dyads in which cases $i$ cites case $j$, and case $j$ cites case $i$) in $C_{\leq t}$. In practice, this is the number of mutual dyads in $C_{\leq t}$, since mutual dyads must emerge among opinions that were written in the same term. The log odds interpretation of the reciprocity statistic is that if opinions in case $i$ cite case $j$, the log odds that case $j$ also cites $i$ change by the theta relative to the configuration in which case $j$ does not cite case $i$.

Accouting for the popularity effect through the in-2-star statistic is prone to cause model degeneracy. In an attempt to stabilize the model against model degeneracy as well as to still capture the popularity effect, \citet{SnijdersTomA.B..2006} introduces the \textit{alternating-in-k-star} statistic, which was shown to be equivalent to the \textit{geometrically weighted in-degree distribution} (gwidegree) statistic introduced by \citet{hunter2006inference}. Define $D_r(C_t, C_{<t})$ as the number of nodes with indegree $r, r \in \{0, \dots , N_t -1 \}$, where $N_t$ is the number of cases in the network at time $t$, then gwidegree represents a network's indegree distribution in a single statistic by geometrically weighting the degree distribution
$$h_{gwidegree}(C_{t}, C_{<t}, \lambda)= \lambda \sum_{r=1}^{N_t-1}\biggl(1-\bigl(1-\frac{1}{\lambda}\bigr)^r\biggr)D_r(C_t, C_{<t}). $$
The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on $D_r(C_t, C_{<t})$ as $r$ increases.
Adding one single edge to the network changes the indegree distribution of the network so that a case with an indegree of $s$ turns into a case with an indegree of $s + 1$. With our notation this means that $D_r$ and $D_{r+1}$ get replaced by
$D_r -1$ and $D_{r+1} +1$, while no changes for $D_\ell,
 \ell \in \{0, \dots , N_t -1 \} \setminus \{r, r-1\} $ are
made. This changes the log odds in the following way
$$\text{logit} P(C_{ij,t}=1 | C_{-ij,t}, C_{<t}) = \exp \bigl(~\theta \cdot( 1- \frac{1}{\lambda})^r \bigr).$$

%In order to account for the popularity effect, we add the \textit{geometrically weighted in-degree distribution} (gwidegree) statistic to the model. The geometrically weighted distribution statistics were introduced by \citet{Hunter.2006} to counteract model degeneracy.

%The log odds interpretation of the triangle statistic is that  $\theta$ gives the change in the log odds of case $i$ citing case $j$ associated  with each additional triangle that would be created by case $i$ citing case $j$ (e.g., if opinions in cases $i$ and $j$ cite one common case $k$, the log odds that case $i$ cites case $j$ change by theta relative to the configuration in which cases $i$ and $j$ do not cite any of the same cases).
